---
title: "MP3"
author: "Tingwei Hu, Brendan Fox, Minho Heo, Christian Peterson"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpmisc)
```

1. What is the stiffness of the intestine in longitudinal and axial loading? How does the stiffness in these two directions compare?
2. What evidence do you have that an intestinal tube is anisotropic?
3. What do you think the function of anisotropy is in the intestine?
4. Did beam orientation result in different flexural stiffness?
5. Did strain rate affect wood flexural stiffness and did this response vary according to how wet the material was?
6. How would this be important in loads applied to a human skeletal element?
7. In your beam analysis, did you find any evidence of viscoelasticity?


# Introduction




# Materials and Methods

## Question 1

Load was calculated in our shared group 3 google sheet using normalized values from our recordings. All formula are in the excel sheet.  

```{r, intestine, message=FALSE}
#all units are SI

#create longitudinal loading data frame
longi.pigdf <- data.frame(
  load.g = c(0.062784,0.125568,0.188352,0.251136,0.31392),
  load.strain = c(0.02666666667,0.05333333333,0.06666666667,0.08,0.1066666667),
  unload.strain = c(0.08,0.09333333333,0.09333333333,0.1066666667,NA),
  area = 0.4/10000
)
#calculate stress and plot stress strain curve
longi.pigdf <- longi.pigdf %>% 
  mutate(stress = load.g/area)
longi.pigdf %>% 
  ggplot(aes(x=load.strain,y=stress))+geom_point()+geom_smooth(method=lm,se=FALSE)+
  stat_poly_eq(aes(label =paste(after_stat(eq.label),after_stat(rr.label), sep = "*\", \"*")))
  
#creating circumferential loading data frame
circum.pigdf <- data.frame(
  load.g = c(6.4,12.8,19.2,25.6,32),
  load.strain = c(0.2222222222,0.4444444444,0.4444444444,0.5555555556,0.6666666667),
  unload.strain = c(0.4444444444,0.5555555556,0.5555555556,0.5555555556,NA),
  area = 1.18/10000
)

#calculate stress and plot stress strain curve
circum.pigdf <- circum.pigdf %>% 
  mutate(stress = load.g/area)
circum.pigdf %>% 
  ggplot(aes(x=load.strain,y=stress))+geom_point()+geom_smooth(method=lm,se=FALSE)+
  stat_poly_eq(aes(label =paste(after_stat(eq.label),after_stat(rr.label), sep = "*\", \"*")))


```
## Question 2

## Question 3

## Question 4


```{r,}
bend.files <- list.files(pattern='*.csv',full.names=FALSE)

benddf <- data.frame(
    dry.wet = NA,
    flat.flip = NA,
    force = NA,
    deflection = NA,
    length = NA,
    flx.stiff = NA
  )

for (i in length(bend.files)){
  metdat <- str_split(bend.files[i], pattern = '_')
  catchdf <- data.frame(
    dry.wet = metdat[[1]][2],
    flat.flip = metdat[[1]][3],
    force = NA,
    deflection = NA,
    length = 35/1000,
    flx.stiff = NA
  )
  tmpdf <- read_csv(bend.files[1]) %>% 
    mutate(
      strain = abs(strain), 
      g = abs(g)
    )
  catchdf$force[1] <- max(tmpdf$g) / 1000 * 9.81
  catchdf$deflection[1] <- 1.5 / 1000
  catchdf$flx.stiff[1] <- (catchdf$force[1] * (catchdf$length[1])^3) / (3 * catchdf$deflection[1])
  #combine all conditions!!
  
  bind_rows(benddf,catchdf)
}

print(benddf)
```

# Results



# Discussion



# Author Contributions

# References