---
title: "MP4"
author: "Nate Mange"
date: "2023-04-12"
output: html_document
bibliography: 'MP4.bib'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
if(!require(ggpubr)) {install.packages(ggpubr)}
library(ggpubr)
if(!require(jpeg)) {install.packages(jpeg)}
library(jpeg)
if(!require(ggpmisc)) {install.packages(ggpmisc)}
library(ggpmisc)
```

# Introduction

MP4 delves into the effects of wing shape on flow patterns with foils moving in a fluid. We aimed to predict bird flight behavior by examining the lift and drag components. We used CAD software and 3D printing technology to create two biomimetic wings for the puffin and the swift that reflect their wing morphologies. Flow patterns were analyzed through particle tracking. We strove to establish the flight characteristics that the wings impart to our birds based on lift, drag, and flow pattern data.


# Methods




# Results

```{r,include=TRUE,message=FALSE,echo=FALSE}
puffin<-read.delim("groupjerry_puffin.txt", header = TRUE, sep = ",",dec = ".")

print(paste0('Mean puffin wing lift: ',as.character(mean(puffin$lift))))
#26.65229
print(paste0('Mean puffin wing drag: ',as.character(mean(puffin$drag))))
#5.229446

Cd_puff <- (mean(puffin$drag))/(0.5*1000*0.1^2*0.00102)
Cl_puff <- (mean(puffin$lift))/(0.5*1000*0.1^2*0.00102)
ClCd_puff <- Cl_puff/Cd_puff


swift<-read.delim("groupjerry_swift.txt", header = TRUE, sep = ",",dec = ".")

print(paste0('Mean swift wing lift: ',as.character(mean(swift$lift))))
#26.65229
print(paste0('Mean swift wing drag: ',as.character(mean(swift$drag))))
#5.229446

Cd_swift <- (mean(swift$drag))/(0.5*1000*0.1^2*0.00102)
Cl_swift <- (mean(swift$lift))/(0.5*1000*0.1^2*0.00102)
ClCd_swift <-Cl_swift/Cd_swift

print(paste0('Lift to drag ratio for puffin: ', as.character(ClCd_puff)))
print(paste0('Lift to drag ratio for swift: ', as.character(ClCd_swift)))

```
```{r,ar vs lift to drag,message=FALSE,echo=FALSE}
class_dat <- read_csv('classdata.csv')

class_dat <- class_dat %>% 
  drop_na() %>%
  rename('l2d' = 'Cl/Cd')

class_dat %>% 
  ggplot(aes(AR,l2d))+geom_point()+geom_smooth(method=lm,se=FALSE)+
  stat_poly_eq()+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())+
  ylab('Lift to Drag Ratio')+
  xlab('Aspect Ratio')+
  ggtitle('Correlating lift to drag ratio to aspect ratio')

lm1 <- lm(l2d ~ AR, class_dat)
lm1_summ <- summary(lm1)
print('Model formula: lm(l2d ~ AR)')
print('Coefficients')
lm1_summ$coefficients
rsqr <- lm1_summ$r.squared
print(paste0('R-squared: ',as.character(rsqr[1])))
confint(lm1)
```

# Discussion


# Author Contributions


# References

